---
title: "tutorial3_recap"
author: "XuanlongQIN"
date: '2024-02-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recap turorial 1 - Basic concepts in R 

### 1.1 Data type
```{r}
name <- c("Bob","Alice","Jack")
name
class(name)
```
### 1.2 Create a dataframe
```{r}
name <- c("Bob","Alice","Jack")
score <- c(80,90,70)
class <- c (1,2,1)
students_scores <- data.frame(names = name,scores = score, classes = class)

class(students_scores)
```

### 1.3 View the data frame
```{r}
students_scores
```
Learn to use `View()` function to check your dataset.

Example: 
```
View(students_scores)
```

### 1.4 Input external dataset
```{r}
# find current path
getwd()


```
Standard methods:

1. put your file in this folder

2. select your own path

Use `head()` check the first 5 rows' data.

### 1.5 Using `help` function
`help()` is a system function which could help you know the syntax and arguments of this function.

Example:
```
help("data.frame")

```


## Recap turorial 2 - igraph in R 
### 2.1 install igraph
Checking whether this package has been installed.
```{r}
library(igraph)
```

### 2.2 Create simple graph
```{r}
g <- graph(c("A","B","B","C","C","D","D","E","E","A","C"),directed = TRUE)
plot(g)
```

### 2.3 Indicators of Networks
#### 2.3.1 Degree Centrality
```{r}
## calculate all vertices
degree_centrality <- degree(g)
degree_centrality

degree_centrality_2 <- degree(g,v = c("A","B"))
degree_centrality_2
```
#### 2.3.2 Betweenness Centrality
```{r}
## calculate all vertices
betweenness_centrality <- betweenness(g)
betweenness_centrality
```

#### 2.3.3 Closeness Centrality
```{r}
## calculate all vertices
closeness_centrality <- closeness(g)
closeness_centrality
```

### 2.4 How to draw a nice picture
Question: How to distinguish color with gender?
What I have known is their gender in a vector.

```{r}
library(igraph)
g <- graph(c("A", "B", "B", "C", "C", "D", "D", "E", "E", "A", "C"), directed = TRUE)
plot(g)
```

```{r}
## Define the gender
### create a gender vector
genders <- c("A" = "male", "B" = "male", "C" = "male", "D" = "female", "E" = "female")
genders




### get the "person" who is male, use names() function apply to all vectors
#genders[genders == "male"]

selected_person_male <- names(genders[genders == "male"])
selected_person_male
```
Using `ifelse` loop assign colors of gender
```{r}
# Assign colors based on gender
vertex_colors <- ifelse(V(g)$name %in% selected_person_male, "blue", "red")
vertex_colors
# Plot the graph
plot(g, vertex.color = vertex_colors)
```

