---
title: "Tutorial_1"
output: html_document
date: "2026-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Learn R basics and syntax

### R Syntax

```{r}
"hello world"
```

```{r}
5

# This is a comment

print ("hello world") # this code is use to print the result.

```

## Work with R variables and data types

### Variable

Variables are containers for storing data values.

```{r}
name <- "John"
age <- 40

name   # output "John"
age    # output 40
```

Print / Output Variables

```{r}
name <- "John"
print(name)
```

And there are times you must use the print() function to output code, for example when working with for loops (which you will learn more about in a later chapter):

```{r}
for (x in 1:10) {
  print(x)
}
```

> **`print()` function is a powerful function in coding, a strong debugging tool.**

### Data types

In programming, data type is an important concept.

Variables can store data of different types, and different types can do different things.

In R, variables do not need to be declared with any particular type, and can even change type after they have been set

```{r}
# numeric
x <- 10.5
class(x)

# integer
x <- 1000L
class(x)

# complex
x <- 9i + 3
class(x)

# character/string
x <- "R is exciting"
class(x)

# logical/boolean
x <- TRUE
class(x)
```

Type conversion

```{r}
x <- 1L # integer
y <- 2 # numeric

# convert from integer to numeric:
a <- as.numeric(x)

# convert from numeric to integer:
b <- as.integer(y)

# print values of x and y
x
y

# print the class name of a and b
class(a)
class(b)
```

Why we need convert types?

In R, you can use operators to perform common mathematical operations on numbers.

```{r}
a <- 4
b <- "hello" # A string is surrounded by either single quotation marks, or **double quotation marks**:


c <- 5.4


d <- a + c

# e <- a + b
print (d)
# print (e)

```

Logistic variable

- In programming, you often need to know if an expression is true or false.

- When you compare two values, the expression is evaluated and R returns the logical answer

```{r}
10 > 9    # TRUE because 10 is greater than 9
10 == 9   # FALSE because 10 is not equal to 9
10 < 9    # FALSE because 10 is greater than 9
```

## 2. R Operators and Control structures (if, while, for)

### R Operators

Arithmetic Operators

```{r}
1+3
```

Assignment Operators

```{r}
my_var <- 3
my_var
```

Comparison Operators

```{r}
a = 5
b = 6

a > b
```

Logical Operators `&` Element-wise Logical AND operator. Returns TRUE if both elements are TRUE `|` Elementwise- Logical OR operator. Returns TRUE if one of the statements is TRUE

```{r}
a = 3
b = 4
a&b
```

### Control structures (If statement)

An "if statement" is written with the if keyword, and it is used to specify a block of code to be executed if a condition is TRUE.

```{r}
a <- 33
b <- 200

if (b > a) {
  print("b is greater than a")
}

if (b < a) {
  print("someting wrong")
}
```

The else if keyword is R's way of saying "if the previous conditions were not true, then try this condition":

```{r}
a <- 33
b <- 33

if (b > a) {
  print("b is greater than a")
} else if (b < a) {
  print ("b is smaller than a")
} else {
  print( "oh! they are equal")
}
```

`or`

The `|` symbol (or) is a logical operator, and is used to combine conditional statements:

```{r}
a <- 200
b <- 33
c <- 500

if (a > b | a > c) {
  print("At least one of the conditions is true")
}
```

### Control structures (while loop)

With the while loop we can execute a set of statements as long as a condition is TRUE:

```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i + 1
}
```

`break` function With the break statement, we can stop the loop even if the while condition is TRUE:

```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i + 1
  if (i == 4) {
    break
  }
}
```

For loop

```{r}
for (x in 1:10) {
  print(x)
}
```

## 3. Create and use functions in R

A function is a block of code which only runs when it is called. - Why we need a function?

Create a function

```{r}
my_function <- function() {
  print("we are good!")
}

my_function ()
```

Arguments

```{r}

my_function <- function (a,b) {
  c = a * b
  return(c)
}

my_function (1,2)

```

## 4. Work with data structures (vectors, lists, matrices, arrays)

Vectors

A vector is the most basic data structure in R. It contains a list of items of the same type.

```{r}
# Vector of strings
fruits <- c("banana", "apple", "orange")

# Print fruits
fruits
```

Lists

A list can hold different types of data in one structure. You can combine numbers, strings, vectors, and even other lists.

```{r}
# List of strings
thislist <- list("apple", "banana", 50, 100)

# Print the list
thislist
```

Matrices

A matrix is a 2D data structure where all elements are of the same type. It is like a table with rows and columns.

```{r}
# Create a matrix
thismatrix <- matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2)

# Print the matrix
thismatrix
```

Arrays

An array is like a matrix but can have more than two dimensions. It stores elements of the same type in multiple dimensions.

```{r}
# An array with one dimension with values ranging from 1 to 24
thisarray <- c(1:24)
thisarray

# An array with more than one dimension
multiarray <- array(thisarray, dim = c(4, 3, 2))
multiarray
```

> Arrays are useful for working with 3D or higher-dimensional data.

Data Frames

A data frame is like a table in a spreadsheet. It can hold different types of data across multiple columns.

```{r}
# Create a data frame
Data_Frame <- data.frame (
  Training = c("Strength", "Stamina", "Other"),
  Pulse = c(100, 150, 120),
  Duration = c(60, 30, 45)
)

# Print the data frame
Data_Frame
```

## 5. Make plots and visualize data (line, scatter, pie, bar)

`plot()` is commonly used in visualization.

```{r}
help ("plot")
```

```{r}
plot (1,3)

plot(c(1, 8), c(3, 10))

plot(c(1, 2, 3, 4, 5), c(3, 7, 8, 9, 12))

plot(1:10, main="My Graph", xlab="The x-axis", ylab="The y axis")

plot(1:10, type="l")

```

A pie chart is a circular graphical view of data.

Use the pie() function to draw pie charts
```{r}
# Create a vector of pies
x <- c(10,20,30,40)

# Display the pie chart
pie(x)
```

Bar Charts
A bar chart uses rectangular bars to visualize data. Bar charts can be displayed horizontally or vertically. The height or length of the bars are proportional to the values they represent.

Use the barplot() function to draw a vertical bar chart:

```{r}
# x-axis values
x <- c("A", "B", "C", "D")

# y-axis values
y <- c(2, 4, 6, 8)

barplot(y, names.arg = x)
```

## 6. Perform basic statistics (mean, median, mode)

Data Set
A data set is a collection of data, often presented in a table.

There is a popular built-in data set in R called "mtcars" (Motor Trend Car Road Tests), which is retrieved from the 1974 Motor Trend US Magazine.

In the examples below (and for the next chapters), we will use the mtcars data set, for statistical purposes.

```{r}
# Step 1. Check the dataset
mtcars
class(mtcars)
```

Get the information of dataset

```{r}
# help (mtcars)
?mtcars
```


Use the dim() function to find the dimensions of the data set, and the names() function to view the names of the variables:

```{r}
Data_Cars <- mtcars # create a variable of the mtcars data set for better organization

# Use dim() to find the dimension of the data set
dim(Data_Cars)

# Use names() to find the names of the variables from the data set
names(Data_Cars)
```


Use the rownames() function to get the name of each row in the first column, which is the name of each car
```{r}
Data_Cars <- mtcars

rownames(Data_Cars)

```

Print Variable Values
If you want to print all values that belong to a variable, access the data frame by using the `$` sign, and the name of the variable (for example mpg (Miles/gallon))

```{r}
get_mpg <- mtcars$mpg
get_mpg
```

Analyzing the data - generally, the first step when you get a data

```{r}
Data_Cars <- mtcars

summary(Data_Cars)

help(summary)
```

R Max and Min
the min() and max() functions can be used to find the lowest or highest value in a set

```{r}
# Find the largest and smallest value of the variable hp (horsepower).
Data_Cars <- mtcars

max(Data_Cars$hp)
min(Data_Cars$hp)

# Also mean() for mean value
mean(Data_Cars$hp)
m = mean(Data_Cars$hp)
cat("mean value is:",m)
print("mean value is:",m) # why print is not available ?
median(Data_Cars$hp)
```

## Hands-on: from data to a plot
Visualize the relationship between vehicle weight and fuel consumption in the `mtcars` dataset;
plotting a regression line for this relations;
color-coding the points by the number of cylinders.

```{r}
mtcars

# Scatter plot
plot(mtcars$wt, mtcars$mpg) # wt	Weight (1000 lbs); Miles/(US) gallon

```

Regression model - Next Tutorial

```{r}
plot(mtcars$wt, mtcars$mpg)
abline(lm(mpg ~ wt, data = mtcars), col = "red") # lm-linear regression
```

Visualization
```{r}
cyl_f <- factor(mtcars$cyl)

plot(mtcars$wt, mtcars$mpg,
     col = cyl_f, # different colors for numbers of cylinders 
     pch = 19) 

legend("topright",
       legend = levels(cyl_f),
       col = 1:3,
       pch = 19,
       title = "Number of cylinder")
```