---
title: "Tutorial_2"
output: html_document
date: "2026-10-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part A · Statistical Models

In Tutorial 1, we described data with `mean()`, `summary()` and plots. Today we ask two further questions:

1. Is a pattern **real**? (hypothesis testing)
2. How does one variable **explain** another? (regression models)

## 1. Hypothesis testing

Hypothesis testing compares two statements:

- **Null hypothesis (H0)**: nothing is going on (no difference, no association)
- **Alternative hypothesis (H1)**: something is going on

We compute a **p-value** — the probability of seeing data like this *if H0 is true*.

- p < 0.05 → reject H0 ("statistically significant")
- p ≥ 0.05 → we cannot reject H0

> A small p-value means "H0 looks unlikely", not "my theory is proven".

### t-test: compare two group means

Question: does fuel efficiency (`mpg`) differ between automatic and manual cars?

```{r}
# am: 0 = automatic, 1 = manual
# convert to a factor with readable labels (see Tutorial 1)
mtcars$am <- factor(mtcars$am,
                    levels = c(0, 1),
                    labels = c("automatic", "manual"))

table(mtcars$am)

# t-test: is the mean mpg different between the two groups?
t.test(mpg ~ am, data = mtcars)
```

Read the output:

- `mean in group automatic` vs `mean in group manual` — the group means
- `t = ...` — the test statistic
- `p-value = ...` — here p < 0.05, so we reject H0: the difference is statistically significant
- In this sample, manual cars are more fuel-efficient

### chi-square test: association between two categorical variables

Question: is the number of cylinders associated with transmission type?

```{r}
# cross table
tbl <- table(mtcars$cyl, mtcars$am)
tbl

# chi-square test: H0 = no association
chisq.test(tbl)
```

Read the output:

- `p-value` < 0.05 → the two variables are associated
- Look at the cross table to tell the direction of the association
- **Warning in the output?** Some expected counts are < 5, so the approximation is unreliable. For small tables R offers an alternative — Fisher's exact test:

```{r}
fisher.test(tbl)
```

## 2. Linear regression

### Simple linear regression: one predictor

A linear model fits the line that best describes the relationship y ~ x:

y = a + b·x + error

We already drew this line in Tutorial 1's hands-on. Now let's read the numbers.

```{r}
model1 <- lm(mpg ~ wt, data = mtcars)
summary(model1)
```

Read the output:

- `(Intercept)` — a: the predicted mpg when wt = 0
- `wt` — b: a one-unit (1,000 lbs) increase in weight is associated with a 5.344 mpg **decrease**
- `Pr(>|t|)` — the p-value of b; here < 0.001, significant
- `Multiple R-squared` — 0.7528: weight alone explains about 75% of the variation in mpg

Visualize the model:

```{r}
plot(mtcars$wt, mtcars$mpg)
abline(model1, col = "red")
```

### Multiple regression: more than one predictor

```{r}
model2 <- lm(mpg ~ wt + hp, data = mtcars)
summary(model2)
```

- With several predictors, b means: the change in y **holding the other variables constant**
- Compare `Multiple R-squared` across models: 0.827 (model2) vs 0.753 (model1) — adding hp helps
- Another criterion: AIC — lower is better

```{r}
AIC(model1, model2)
```

### Categorical predictors

Categorical variables must be factors (Tutorial 1):

```{r}
model3 <- lm(mpg ~ wt + factor(cyl), data = mtcars)
summary(model3)
```

- `factor(cyl)6` and `factor(cyl)8` compare against the reference group (4 cylinders)
- 8-cylinder cars: 4.26 mpg lower than 4-cylinder cars, holding weight constant

## 3. Logistic regression

### When the outcome is binary

Digital sociology examples: retweet or not, follow back or not, purchase or not.

Why not `lm()`? A straight line can predict values outside [0, 1] — meaningless for a probability. Logistic regression keeps predictions inside (0, 1).

```{r}
# am is now a factor, so the model predicts P(am = "manual")
model4 <- glm(am ~ wt + hp, data = mtcars, family = binomial)
summary(model4)
```

### Interpretation: odds ratios

The coefficients are in **log-odds** — hard to read directly. Convert them with `exp()`:

```{r}
exp(coef(model4))
```

Read the output:

- `wt` odds ratio = 0.04: one unit more weight → odds of manual transmission × 0.04 (much less likely)
- `hp` odds ratio = 1.05: one more hp → odds × 1.05
- OR = 1 → no effect; OR > 1 → positive association; OR < 1 → negative association

> Rule of thumb: look at your **outcome variable** first, then choose the model.

# Part B · Network Analysis

## 1. What is a network?

A network is a set of **nodes** (vertices) connected by **edges**.

- Nodes — people, accounts, organizations
- Edges — ties between them (follows, retweets, friendships)

Types of networks:

- **Directed** — A follows B ≠ B follows A (Weibo/Twitter)
- **Undirected** — friendship is mutual
- **Weighted** — number of messages, strength of the tie

The most common data format: an **edge list** — one row per tie, with a `from` and a `to` column.

## 2. Build and plot a graph with igraph

```{r}
# install.packages("igraph")   # run once if not installed
library(igraph)
```

Our first example: **Zachary's karate club** (1977). 34 members of a university karate club; after a conflict between the instructor and the administrator, the club split into two groups. The network is built into igraph.

```{r}
g <- make_graph("Zachary")
g

# nodes and edges
V(g)
E(g)

# how big is the network?
vcount(g)
ecount(g)
```

```{r}
plot(g)
```

## 3. Degree & density

**Degree** — how many ties a node has.

- in-degree: followers; out-degree: following
- in digital sociology, the influencer = high in-degree

```{r}
deg <- degree(g)
deg

# who has the most ties?
V(g)$name[which.max(deg)]
```

Degree distribution:

```{r}
hist(degree(g), main = "Degree distribution", xlab = "Degree")
```

**Density** — how connected the network is overall (0 to 1):

```{r}
edge_density(g)
```

## 4. Centrality: who is important?

Different measures capture different kinds of importance:

- **Degree centrality** — most ties (popular, active)
- **Betweenness centrality** — lies on many shortest paths (bridge, gatekeeper)
- **Closeness centrality** — close to everyone else (information spreads fast)

```{r}
bet <- betweenness(g)
clo <- closeness(g)

# top-3 by each measure
V(g)$name[order(bet, decreasing = TRUE)[1:3]]   # by betweenness
V(g)$name[order(clo, decreasing = TRUE)[1:3]]   # by closeness
```

> The same person can be "important" in one way and not another — choose the measure that fits your research question.

## 5. Community detection

Find groups that are densely connected **within** and sparsely connected **between**.

The karate club really split into two clubs — can the algorithm find this split from the network alone?

```{r}
cm <- cluster_fast_greedy(g)
membership(cm)

# how many communities, how big?
sizes(cm)

# plot with communities highlighted
plot(cm, g)
```

- The algorithm recovers the factions using only the network structure, no attribute data
- Check the two leaders: the instructor is node 1 ("Mr Hi"), the administrator is node 34 ("John A")

```{r}
V(g)$name[c(1, 34)]
```

## 6. A social media example

Now a directed follower network of 24 user accounts. The data is in `social_edges.csv` (same folder as this file) — an edge list where each row means "from follows to".

```{r}
edges <- read.csv("social_edges.csv")
head(edges)

sg <- graph_from_data_frame(edges, directed = TRUE)
sg

plot(sg, vertex.size = 8, vertex.label.cex = 0.7)
```

Find the influencers — top in-degree (followers):

```{r}
in_deg <- degree(sg, mode = "in")
sort(in_deg, decreasing = TRUE)

# top-3 influencers
V(sg)$name[order(in_deg, decreasing = TRUE)[1:3]]
```

Community structure of the follower network:

```{r}
# cluster_fast_greedy works on undirected graphs only.
# For a directed graph, use walktrap (random walks) or infomap instead.
cm2 <- cluster_walktrap(sg)
sizes(cm2)
plot(cm2, sg, vertex.size = 8, vertex.label.cex = 0.7)
```

> `igraph` works the same way for 34 karate members and 34 million users.

## Hands-on

**Task A · Statistical models** (`mtcars`):

1. t-test — does fuel efficiency differ between automatic and manual cars?
2. Linear regression — model `mpg` with `wt + hp`; interpret the coefficients and R²
3. Logistic regression — model `am`; compute and interpret the odds ratios

**Task B · Network analysis:**

1. Build and plot the karate club network
2. Compute degree and betweenness — who are the top-3 most central?
3. Run community detection — compare with the real split
4. Load `social_edges.csv` — find the top influencers and plot the network

**Do:** the hands-on exercises end-to-end on your own machine.
